Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance in zooplankton

### 2022-06-06 text/Manuscript.Rmd
Figs ms updates

![**Figure 1. Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (A and B) Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (magenta). Co-staining with acetylated α-tubulin antibody as marker for cilia and axonal scaffold (ac α-tub: green). INNOSs with the entire cell body and axon labelled are shown with white arrowheads and with only the cell body labelled are shown with black arrowheads. Views of larvae at 2-day-old (A) and 3-day-old (B) from the anterior side, respectively. (C) Three types of neurons (cPRC, INNOS and INRGWa) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva. The INNOS on the ventral side labelled with (A) and (B) are shown in purple. Other neurons are shown in light grey. Axons and dendrites appear as lines and cell body positions are represented by spheres. (D and E) Expression analysis of the NOS gene (magenta) using in situ HCR. Anterior view of the larva at 2-day-old. (D) Co-stained image using acetylated α-tubulin antibody (ac α-tub: green). (E) Co-expression analysis image with the RYamide proneuropeptide gene (RYa-pNP: green), which is highly expressed in INNOS. White arrowheads indicate co-expressed INNOS. Black arrowheads indicate larval eyes. White dashed lines indicate the outer frame of the larva. (F) Wiring diagram of the cPRC circuit. Hexagons represent neuronal groups (numbers indicated in square brackets), arrows represent synaptic connections (numbers indicated in number of synapses); NOS and RGW indicate INNOS and INRGW-dc1 respectively. Ciliomotor neuron shows two serotonergic neurons (Ser-h1).](../figures/Figure1.png){width="100%"}

@@

![**Figure 2. Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** (A) Percentage of time in which cilia are closed in DMSO-treated larvae (control) vs. larvae exposed to NOS inhibitor. (B) Top: The domain structure of NOS. Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target sites. Yellow shows PAM sequences, red shows stop codons. (C and D) Vertical swimming in wild type and mutant larvae at 2-day-old (C) and 3-day-old (D) stimulated with blue (480 nm) and UV (395 nm) light from above. the data are shown in 30 s bins.](../figures/Figure2.png){width="100%"}

@@

![**Figure 3. cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** (A) Phylogenetic tree of guanylate cyclase by maximum likelihood (ML). Guanylate cyclase-coupled receptor and soluble guanylate cyclase as outgroups. Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata. Tree robustness was tested with 1000 replicates of ultrafast bootstrap. Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum). (B-D) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR. White arrowheads indicate cells expressing NIT-GC1. (B) Co-staining image using acetylated α-tubulin antibody (ac α-tub: green). (C) Co-expression analysis image of the c-opsin1 gene (green). Black arrowheads indicate cells expressing c-opsin1 differently from cPRC. (D) Co-expression analysis image of the MLD-pedal2 amide proneuropeptide gene (MLD-pNP: green). Black arrowheads indicate cells different from cPRC. The white dashed lines in (C) and (D) indicate the outer frame of the larvae. (E, F) Localisation analysis using NIT-GC1 and NIT-GC2 antibodies. Green shows co-staining with acetylated α-tubulin antibody (ac α-tub). (G) Co-localisation analysis using NIT-GC1 and NOS antibodies. (B-G) Anterior view of the 2-day-old larvae.](../figures/Figure3.png){width="100%"}

@@

![**Figure 4. NIT-GC1 is induced by NO to produce cGMP.** (A) Top: The domain structure of NIT-GC1. Grey squares indicate the transmembrane domain. Bottom shows a schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control. Green cGull shows cGMP-dependent green fluorescence. (B) The intensity changes in cGMP fluorescence over time for the three conditions are shown. Gray traces show the optical responses to individual intensities (n = 6). The individual intensities are normalized (ΔF/F0). The dark line shows the mean value and the band the standard error range of the mean. The respective solution was acted on 2 min after the start of imaging (grey bars). (C) The mean intensity (Fmean/F0) after acting on the solution. Multiple comparisons were made using the Tukey honestly significant difference test.](../figures/Figure4.png){width="100%"}

@@

![**Figure 5. NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) An INNOS visualised by DAF-FM is shown (a white arrowhead). A black arrowhead indicates the position of the adult eye. The white dashed line indicates the outer frame of the larva. All subsequent images are of 2-day-old larvae. (B) Dual-colour imaging of DAF-FM (green) and RGECO1a (magenta) allows simultaneous visualisation of the INNOS and cPRCs. (C) The intensity of DAF-FM in INNOS changes over time when cPRC (top) and other cells (NC) (bottom) are stimulated with UV (405 nm). Hereafter, light purple squares indicate the timing of UV stimulation, gray traces show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark lines and the bands show the mean value and the standard error range of the mean. (D) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (purple) and NOS mutant (NOSΔ23) (gray) larvae. (E) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 knockdown by morpholino (NITGC1-MO1 and 2) larvae. (F) Co-expression analysis image of the cPRC marker gene (NIT-GC1: magenta) and the INNOS marker gene (RYa-pNP: green) using in situ HCR. White arrowheads indicate INNOS strongly expressing RYa-pNP. (G) INNOS (blue dashed line) identified from calcium imaging using GCaMP6s by their position in relation to the cPRC (purple dashed line). (H) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INNOS (blue) when cPRC is stimulated with UV in WT (left) and NOS mutant (NOSΔ11: right) larvae. (I) Schematic diagram of the immunostaining procedure after calcium imaging. (J) The position of the cPRC (purple dashed line) in calcium imaging with GCaMP6s. Cells around the cPRC are numbered according to the position of the nuclei. (K) Immunostaining image using RGWamide (RGWa) and serotonin (Ser) antibodies (red) and DAPI (light blue) in the same region as (J). INRGW (green dashed line) and Ser-h1 (black arrowheads) identified from the numbered nuclear positions are shown. (L) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INRGW (green) when the cPRC are stimulated with UV in WT (left) and NOS mutant (NOSΔ23: right) larvae.](../figures/Figure5.png){width="100%"}

>>>>

![**Figure 1. Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (A and B) Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (magenta). Co-staining with acetylated α-tubulin antibody as marker for cilia and axonal scaffold (ac α-tub: green). INNOSs with the entire cell body and axon labelled are shown with white arrowheads and with only the cell body labelled are shown with black arrowheads. Views of larvae at 2-day-old (A) and 3-day-old (B) from the anterior side, respectively. (C) Three types of neurons (cPRC, INNOS and INRGWa) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva. The INNOS on the ventral side labelled with (A) and (B) are shown in purple. Other neurons are shown in light grey. Axons and dendrites appear as lines and cell body positions are represented by spheres. (D and E) Expression analysis of the NOS gene (magenta) using in situ HCR. Anterior view of the larva at 2-day-old. (D) Co-stained image using acetylated α-tubulin antibody (ac α-tub: green). (E) Co-expression analysis image with the RYamide proneuropeptide gene (RYa-pNP: green), which is highly expressed in INNOS. White arrowheads indicate co-expressed INNOS. Black arrowheads indicate larval eyes. White dashed lines indicate the outer frame of the larva. (F) Wiring diagram of the cPRC circuit. Hexagons represent neuronal groups (numbers indicated in square brackets), arrows represent synaptic connections (numbers indicated in number of synapses); NOS and RGW indicate INNOS and INRGW-dc1 respectively. Ciliomotor neuron shows two serotonergic neurons (Ser-h1).](../figures/Fig1.png){width="100%"}

@@

![**Figure 2. Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** (A) Percentage of time in which cilia are closed in DMSO-treated larvae (control) vs. larvae exposed to NOS inhibitor. (B) Top: The domain structure of NOS. Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target sites. Yellow shows PAM sequences, red shows stop codons. (C and D) Vertical swimming in wild type and mutant larvae at 2-day-old (C) and 3-day-old (D) stimulated with blue (480 nm) and UV (395 nm) light from above. the data are shown in 30 s bins.](../figures/Fig2.png){width="100%"}

@@

![**Figure 3. cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** (A) Phylogenetic tree of guanylate cyclase by maximum likelihood (ML). Guanylate cyclase-coupled receptor and soluble guanylate cyclase as outgroups. Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata. Tree robustness was tested with 1000 replicates of ultrafast bootstrap. Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum). (B-D) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR. White arrowheads indicate cells expressing NIT-GC1. (B) Co-staining image using acetylated α-tubulin antibody (ac α-tub: green). (C) Co-expression analysis image of the c-opsin1 gene (green). Black arrowheads indicate cells expressing c-opsin1 differently from cPRC. (D) Co-expression analysis image of the MLD-pedal2 amide proneuropeptide gene (MLD-pNP: green). Black arrowheads indicate cells different from cPRC. The white dashed lines in (C) and (D) indicate the outer frame of the larvae. (E, F) Localisation analysis using NIT-GC1 and NIT-GC2 antibodies. Green shows co-staining with acetylated α-tubulin antibody (ac α-tub). (G) Co-localisation analysis using NIT-GC1 and NOS antibodies. (B-G) Anterior view of the 2-day-old larvae.](../figures/Fig3.png){width="100%"}

@@

![**Figure 4. NIT-GC1 is induced by NO to produce cGMP.** (A) Top: The domain structure of NIT-GC1. Grey squares indicate the transmembrane domain. Bottom shows a schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control. Green cGull shows cGMP-dependent green fluorescence. (B) The intensity changes in cGMP fluorescence over time for the three conditions are shown. Gray traces show the optical responses to individual intensities (n = 6). The individual intensities are normalized (ΔF/F0). The dark line shows the mean value and the band the standard error range of the mean. The respective solution was acted on 2 min after the start of imaging (grey bars). (C) The mean intensity (Fmean/F0) after acting on the solution. Multiple comparisons were made using the Tukey honestly significant difference test.](../figures/Fig4.png){width="100%"}

@@

![**Figure 5. NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) An INNOS visualised by DAF-FM is shown (a white arrowhead). A black arrowhead indicates the position of the adult eye. The white dashed line indicates the outer frame of the larva. All subsequent images are of 2-day-old larvae. (B) Dual-colour imaging of DAF-FM (green) and RGECO1a (magenta) allows simultaneous visualisation of the INNOS and cPRCs. (C) The intensity of DAF-FM in INNOS changes over time when cPRC (top) and other cells (NC) (bottom) are stimulated with UV (405 nm). Hereafter, light purple squares indicate the timing of UV stimulation, gray traces show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark lines and the bands show the mean value and the standard error range of the mean. (D) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (purple) and NOS mutant (NOSΔ23) (gray) larvae. (E) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 knockdown by morpholino (NITGC1-MO1 and 2) larvae. (F) Co-expression analysis image of the cPRC marker gene (NIT-GC1: magenta) and the INNOS marker gene (RYa-pNP: green) using in situ HCR. White arrowheads indicate INNOS strongly expressing RYa-pNP. (G) INNOS (blue dashed line) identified from calcium imaging using GCaMP6s by their position in relation to the cPRC (purple dashed line). (H) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INNOS (blue) when cPRC is stimulated with UV in WT (left) and NOS mutant (NOSΔ11: right) larvae. (I) Schematic diagram of the immunostaining procedure after calcium imaging. (J) The position of the cPRC (purple dashed line) in calcium imaging with GCaMP6s. Cells around the cPRC are numbered according to the position of the nuclei. (K) Immunostaining image using RGWamide (RGWa) and serotonin (Ser) antibodies (red) and DAPI (light blue) in the same region as (J). INRGW (green dashed line) and Ser-h1 (black arrowheads) identified from the numbered nuclear positions are shown. (L) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INRGW (green) when the cPRC are stimulated with UV in WT (left) and NOS mutant (NOSΔ23: right) larvae.](../figures/Fig5.png){width="100%"}
